Clustering galaxies at 0.5 < $z$ < 1: an unsupervised approach

### 2016-05-30 main.tex
after grammatical corrections

\begin{abstract} %Most of this abstract is just a filler % PB160426: OK, no comments on it for now.
In this paper we introduce a new method of clustering high red shift galaxies, based on their SED.
We use unsupervised self organizing map to train and validate an artificial neural network which can categorize the spectral energy distribution (SED) of galaxies. 
We utilize the spectral model introduced by Kinney et al. as a training set and a sample of 142 galaxies with 0.5 < $z$ < 1 a validation set.
We used physical properties of the sample galaxies measured using {\em CIGALE} code to study and compare the properties of each cluster.
In this paper we used an unsupervised self organizing map (SOM)method, to create a network based in SED models of galaxies and used sample of galaxies to test the networks. 
We classify galaxies in different groups and studied the correlation between properties of the galaxies.
we conclude that the SOM is a very useful and accurate method to be used in classifying galaxies base on their morphology.

%%%%%%From this point it is just repeated%%%%%%%%%

\end{abstract}
\begin{keywords} 
 galaxies: star formation, galaxies: stellar content, galaxies: ISM, methods: observational, methods: statistical, data mining, methods:data analysis
\end{keywords}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Intro
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\import{sections0.01/}{intro0.01.tex}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%DATA
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\import{sections0.01/}{data0.01.tex}

@@

\import{sections0.01/}{method0.01.tex}

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Results
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\import{sections0.01/}{results0.01.tex}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Discussion
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Summery
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{SUMMARY}
\label{sec: summary} %PB160427: not so sure I like the numbering of every sentence, or starting almost all of them with "We". But this is a style thing that we can discuss later.
1) We introduced the self organizing maps to be used to categorizing the SED of galaxies.
\newline
2) Using this method we not only can categorized the galaxies with more precision but also we can see the relative similarity or dissimilarity between each type of galaxies.  
\newline
3) We show that using SOM, we could recognize the galaxies have similarity to two or more morphological types of galaxies.
\newline
4) We trained networks using sample of SEDs and test them with sample of galaxies with SEDs in the similar badpasses.
\newline
5) We plot the properties of the galaxies using new classification and see the correlation between age and SSFR and stellar mass of the galaxies.
\newline
6) We show that the properties of the galaxies in each groups is in good agreement between their morphological types.
\newline
7) We show that SOM can help us to easily identified new type of the galaxies, in large surveys very easy. 
\newline
8) This method can be used to find outliers very fast.
\newline
%PB160427: Um, I'm not so sure you really demonstrated numbers 7 and 8. Those are interesting and it could be that a little more work on this and addition to the results section would be a good idea.

\section*{ACKNOWLEDGMENTS}
S.R. and P.B. acknowledge research support from the Natural Sciences and Engineering Research Council of Canada. This research has made use of the NASA/IPAC Extragalactic Database (NED), which is operated by the Jet Propulsion Laboratory, California Institute of Technology, under contract with the National Aeronautics and Space Administration.
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%biblio
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\bibliographystyle{mnras}
\bibliography{ref_mining_h.bib}

>>>>

\begin{abstract}%% (Must be 250 Words)
Spectral energy distribution (SED) of a galaxy contains information about the physical properties of it.
Classifying SEDs using temples helps understand the nature of the source of SEDs.
However, available SED templates are limited, which makes finding type of SEDs tricky.
Most of the current classifying methods cannot classify SEDs outside the ones in the template, and will fail in the special cases.
In this paper, we introduced a new method of clustering high red-shift galaxies, based on their spectral energy distributions (SED)s.
SOM is a type of the artificial neural networks, which mostly is used to identify subgroups in a sample.
We utilized an unsupervised self-organizing map (SOM) as a method to train networks using templates of SEDs ranging from far ultraviolet (FUV) to near infrared (NIR) wavelengths.
Then, we used the trained networks to cluster a sample of 142 galaxies with 0.5 < $z$ < 1.
Since, some the SEDs of these galaxies are not exactly the same as the ones in the template, the previous attempts to classifying SEDs of this group of galaxies were not completely successful. 
By using SOM, we classify those special SEDs and show that they have similarities to two more of the template ones.
We investigated the vapidity of our results, by studying the SEDs, and some physical properties of the galaxies that are averaged over galaxies in each neuron the SOMs.

%%% what is the problem we are trying to solve in general! why do we want o classify thing in general, and what is wrong in exciting classification scheme

\end{abstract}
\begin{keywords} 
 galaxies: high red shifts, galaxies: spectral energy distribution, methods: observational, methods: statistical, data mining, methods:data analysis
\end{keywords}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Intro
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\import{sections0.03/}{intro0.03.tex}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%DATA
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\import{sections0.03/}{data0.03.tex}

@@

\import{sections0.03/}{method0.03.tex}

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Results
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\import{sections0.03/}{results0.03.tex}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Discussion
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Summery
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{SUMMARY AND FUTURE APPLICATIONS}
\label{sec: summary}

% Self organizing maps can be used to classify celestial objects (i.e. stars, quasars, spectra of galaxies, light curves, etc.).
% We showed that we can create various networks with different sizes/dimensions based on information we need from the data. 
% If a broad and general classification is required, networks can have one dimension with few number of neurons.
% While if we need classifications based on differences in more details, we can use a higher number of neurons.
% Since the SOM codes does not work with uncertainty of input parameter, sometimes too much attention to details can cause problems in classifications. 
% Some small differences could easily be the result of atmosphere fluctuation or other instrumental or observational problems. 
% for using SOM method, one should be careful about these small differences and consider in order to separate two groups from each other, how much details do they need to pay attention to.

% We used SOMs to classify SEDs of the galaxies with known morphological type from K96, and created networks with different usages.
% By varying size of networks, we found the relative similarity or dissimilarity between each type of galaxies.
% since we needed a 1D network with 22 neurons to be able to separate all the 12 types of galaxies from K96 for the first time, we concluded that types B and E, and types SB1 and SB2 galaxies are really similar to each other.
% We also showed that networks generated with SOM method can be used to easily identified new type of the galaxies or outlier, in large surveys.

% The sample of 142 high red-shift galaxies from T12 was used to test the trained networks.
% One of the main criteria to chose a test sample is that the test sample must have fluxes in exactly the same wavelengths of the training sample.
% The test results showed that using SOMs can lead us to identify galaxies with SEDs similar to two or more morphological types but not exactly the same as one of them.
% A freedom of having in between types is one of the main differences between supervised and unsupervised ANNs.
% T12  have used a supervised training method, and trained networks with K96 SED template.
% Same as this project, they tried to classify the sample of 142 galaxies using the trained networks.
% However, they could not classify 37 out of 142 galaxies in the sample.
% The unsupervised method that we used in this project, gave us the ability to classify galaxies either as one of the morphological types that have been introduced by K96 or between those types. 
% Therefore, we classified all 142 galaxies of our test sample.
% We plotted the properties of the galaxies in test sample using new classification and found a tighter correlations between mean values of age, sSFR, stellar mass, and extinction in FUV band of the galaxies in new classifications than previous studies.
% We also showed he properties of the galaxies in each groups is in good agreement between their morphological types.

Self-organizing maps can be used to classify celestial objects (e.g. stars, quasars, spectra of galaxies, light curves, etc.).
We showed that we can create various networks with different sizes/dimensions based on the information we need from the data. 
If a broad and general classification is required, networks can have one dimension with a few number of neurons.
If one needs more detailed classifications, a higher number of neurons should be used.
Since the SOM codes does not include the uncertainty of input parameter, sometimes too much attention to details can cause problems in classifications. 
Some small differences could easily be the result of atmosphere fluctuation or other instrumental or observational problems. 
When using SOM method, one should be careful about these small differences and consider them in order to separate two groups from each other.

We used SOMs to classify SEDs of the galaxies with known morphological type from K96, and created networks with different usages.
By varying the size of the networks, we found the relative similarity or dissimilarity between each type of galaxies.
Since we needed a 1D network with 22 neurons to be able to separate all 12 types of galaxies from K96 for the first time, we concluded that types B and E, and types SB1 and SB2 galaxies are very similar to each other.
We also showed that networks generated with SOM method can be used to easily identified new type of the galaxies or outliers in large surveys.

A sample of 142 high red-shift galaxies from T12 was used to test the trained networks.
One of the main criteria to chose a test sample is that the test sample must have fluxes in exactly the same wavelengths of the training sample.
The test results showed that using SOMs can lead us to identify galaxies with SEDs similar to two or more morphological types but not exactly the same as one of them.
The freedom of having in-between types is one of the main differences between supervised and unsupervised ANNs.
T12 have used a supervised training method, and trained networks with K96 SED template.
Same as this work, they tried to classify the sample of 142 galaxies using the trained networks.
However, they could not classify 37 out of 142 galaxies in the sample.
The unsupervised method, which we have used in this project, gave us the ability to classify galaxies either as one of the morphological types that have been introduced by K96 or in between those types. 
Therefore, we classified all 142 galaxies of our test sample.
We plotted the properties of the galaxies in test sample using the new classification and found a tighter correlations between mean values of age, sSFR, stellar mass, and extinction in FUV band of the galaxies in new classifications than previous studies.
We also showed that the properties of the galaxies in each groups is in good agreement between their morphological types.

\section*{ACKNOWLEDGMENTS}
S.R. and P.B. acknowledge research support from the Natural Sciences and Engineering Research Council of Canada. 
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%biblio
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\bibliographystyle{mnras}
\bibliography{ref_mining_h.bib}
